Comparing MTurk and the US Population's Occupational Diversity

### 2018-11-27 Manuscript.Rmd

#- - - - - - - - - - - - - 
#insert table 2 about here
#- - - - - - - - - - - - -

##Replication Demographic Analaysis
##Age Analysis
```{r}
#Age Analysis T- Test
#Load Dataset
require(dplyr)
require(readxl)
age_demo <- read_excel("age_demo.xlsx")

@@

##Gender Analysis
```{r}
##Gender Analysis
BLS_data <- read_excel("employment_demo.xlsx")

@@

##Ethnicity Analysis
```{r}
#Ethnicity Analysis 
#Define obs_eth
BLS_cauc <- as.numeric(BLS_data[7, 4])/100
BLS_AfAm <- as.numeric(BLS_data[7,5])/100
BLS_Asn <- as.numeric(BLS_data[7,6])/100
BLS_Hisp <- as.numeric(BLS_data[7,7])/100
BLS_eth <- c(BLS_AfAm, BLS_Asn, BLS_cauc, BLS_Hisp)

@@

###Education Analysis 
```{r}
#Education Analysis 
#Clean BLS Data. Get totals.  
BLS_education_original <- read_excel("education_demo.xlsx")
BLS_education <- as.data.frame(BLS_education_original[8:14, ])
BLS_education <- BLS_education[1, c(-2, -3, -6)]
names(BLS_education) <- c("category", "High school graduates, no college(1)", "Some college, no degree", "Associate degree", "Bachelor's degree only", "Advanced degree")

@@

#Function to find proportion for each iteration of the vector. 
e_prop <- function(x) {
  e <- x/BLS_total
}

#Use vapply and the above function to get expected proportions from BLS data
BLS_edu <- sapply(BLS_education, e_prop)

>>>>

#- - - - - - - - - - - - - 
#insert table 2 about here
#- - - - - - - - - - - - - 
##Barplot 1
```{r}
#Bar Plot with error bars
require(ggplot2)
Values <- c(Table2[1:13, 5], Table2[1:13, 2])
Group <- c("MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "MTurk", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS")
Occupations <- c(Table2[1:13, 1], Table2[1:13, 1])
Occ_Nums <- sapply(strsplit(Occupations, " "), "[[", 1)
barplot_data <- data.frame(as.character(Values), as.character(Occ_Nums),as.character(Group))
names(barplot_data) <- c("Frequency", "Occupations","Group")
dodge <- position_dodge(width = .9)

#Define Error Bar limits.
upperM <- as.vector(Table2[1:13, 6])
upperB <- c(0, 0,0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
upper <-c(upperM, upperB)

lowerM <- as.vector(Table2[1:13, 6])
lowerB <- c(0, 0,0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
lower <- c(lowerM, lowerB)

se <- data.frame(upper, lower)
limits <- aes(ymax = se$upper, ymin = se$lower)

require(ggplot2)
ggplot(barplot_data, aes(Occ_Nums, Values, fill = Group)) + 
  geom_errorbar(limits, position = dodge,
                width = 0.10, size = 1.0) +
  geom_col(position = dodge)  +
  ggtitle("MTurk's Occupational Diveristy with Error Bars") +
  labs(x = "Occupation", y = "Frequency") +
  theme(legend.position = "top", plot.title = element_text(hjust = 0.5), 
        text=element_text(family="serif"))+
  coord_flip()
```
##2015 Demographic Analysis 
##Study 1 Age
```{r}
#Chris's 2015 age analysis 
require(dplyr)

#Load BLS Data
BLS_age_15 <- read_excel("BLS.Age.2015.xlsx")

#Define population mean. In this case it was already calculated in the dataset. 
pop_mean_15 <- as.numeric(BLS_age_15[6, 10])

#Define sample mean
samp_mean_15 <- as.numeric(MTurk15 %>%
                        summarise(MTurk_age = mean(age, na.rm = TRUE)))

#Sample standard deviation.
samp_sd_15 <- as.numeric(MTurk15 %>%
                        summarise(sd_age = sd(age, na.rm = TRUE)))

#Define sample size
n_15 <- nrow(MTurk15)

#Run t test.
t_value_15 <- (samp_mean_15 - pop_mean_15)/(samp_sd_15/sqrt(n_15))

##Study 1 Gender
```{r}
#Gender Analysis 2015
library(readxl)
BLS_2015_doc <- read_excel("bls.gender.2015.xlsx")

#Get BLS gender data (expected value)
BLS_gender_15_data <- BLS_2015_doc[87:93,c(1,8,22, 28)]
BLS_men_15 <- as.numeric(BLS_gender_15_data[6,3])/as.numeric(BLS_gender_15_data[5,3])
BLS_women_15 <- as.numeric(BLS_gender_15_data[7,3])/as.numeric(BLS_gender_15_data[5,3])
BLS_gender_15 <- c(BLS_women_15,BLS_men_15)
names(BLS_gender_15) <- c("Female", "Male")

#Get MTurk gender data (observed value)
require(dplyr)
MTurk_gender_15 <- MTurk15 %>%
  count(gender)
MTurk_gender_15 <- MTurk_gender_15[2:3,]

#Define confidence intervals
require(binom)
gCI_15 <- binom.confint(x = MTurk_gender_15$n, n = sum(MTurk_gender_15$n), conf.level = 0.95, methods = "exact")
gCI_15 <- gCI_15[ ,4:6]

#Build Gender table with gender, expected value, n, LL, observed, UL
gender_table_15 <- cbind(MTurk_gender_15, gCI_15, BLS_gender_15)
names(gender_table_15) <- c("gender", "n", "o", "LL_g", "UL_g", "e")
gender_table_15 <- gender_table_15[, c(1, 3, 2, 4, 6, 5)]

gender_ch2_15 <- as.data.frame(chi2(gender_table_15))
names(gender_ch2_15) <- c("chi2")
gender_ch2_15 <- na.omit(gender_ch2_15)

#Sum chi-square values.
sum(gender_ch2_15$chi2, na.rm=TRUE)
```

##Study 1 Ethnicity
```{r}
#Ethnicity Analysis 
#Define obs_eth
BLS_ethnicity_15_data <- BLS_2015_doc[c(94, 97, 100), (c(1, 22))]
names(BLS_ethnicity_15_data) <- c("ethnicity", "population")
BLS_ethnicity_15_data$population <- as.numeric(BLS_ethnicity_15_data$population)
BLS15_eth_prop <- sapply(BLS_ethnicity_15_data$population, function(x) {x/sum(BLS_ethnicity_15_data$population)})
names(BLS15_eth_prop) <- c(BLS_ethnicity_15_data$ethnicity)

#Get MTurk ethnicity data
MTurk_eth_15 <- count(MTurk15, ethnicity)

MTurk_eth_15 <- MTurk_eth_15[c(4, 2, 3), ]

#Define confidence intervals 
ethCI_15 <- binom.confint(x = MTurk_eth_15$n, n = sum(MTurk_eth_15$n), conf.level = 0/95, methods = "exact")
ethCI_15 <- ethCI_15[ , 4:6]

#Build Ethnicity table 
ethnicity_table_15 <- cbind(MTurk_eth_15, ethCI_15, BLS15_eth_prop)
names(ethnicity_table_15) <- c("ethnicity", "n", "o", "LL", "UL", "e")
eth_table_15 <- ethnicity_table_15[ , c(1, 6, 2, 4, 3, 5)]

#Ethnicity chi squared test
eth_ch2_15 <- as.data.frame(chi2(eth_table_15))
names(eth_ch2_15) <- c("chi2")
row.names(eth_ch2_15) <- c(BLS_ethnicity_15_data$ethnicity)
eth_ch2_15 <- na.omit(eth_ch2_15)

#Sum chi-square values.
sum(eth_ch2_15$chi2, na.rm=TRUE)
```

##Study 1 Education Level 
```{r}
#2015 Education Analysis 
#Clean BLS Data. Data is split between men and women, so the first step is to total these values for the civilian noninstitutionalize population. 
BLS_education_15_data <- BLS_2015_doc[587:610,c(1,13,24,37,51,61,67)]
BLS_education_15_total <- as.numeric(BLS_education_15_data[4,2]) + as.numeric(BLS_education_15_data[20,2])

#Combine categories as necessary.Add men and women together.
less_than_highschoo_15 <- as.numeric(BLS_education_15_data[4,3]) + as.numeric(BLS_education_15_data[20,3])

#high school/ged includes "some college, no degree"
highschool_ged_15 <- as.numeric(BLS_education_15_data[4,4]) + as.numeric(BLS_education_15_data[20,4]) + as.numeric(BLS_education_15_data[4,5]) + as.numeric(BLS_education_15_data[20,5])

two_year_deg_15 <- as.numeric(BLS_education_15_data[4,6]) + as.numeric(BLS_education_15_data[20,6])

bachelors_or_higher_15 <- as.numeric(BLS_education_15_data[4,7]) + as.numeric(BLS_education_15_data[20,7])

BLS_education_15 <- c(less_than_highschoo_15,highschool_ged_15, two_year_deg_15, bachelors_or_higher_15)
names(BLS_education_15) <- c("less than high school degree", "high school/GED" , "2 year degree", "bachelor's or more")

#Use vapply and the anonomous function to get expected proportions from BLS data
BLS_edu_15 <- sapply(BLS_education_15, function(x){x/BLS_education_15_total})

#Clean MTurk Data
MTurk_education_15 <- MTurk15 %>%
  count(MTurk15$`education level`)
colnames(MTurk_education_15) <- c("education level", "n")
advanced_deg_15 <- as.numeric(MTurk_education_15[6,2]) + as.numeric(MTurk_education_15[8,2]) + as.numeric(MTurk_education_15[3,2])
no_highschool_dip <- as.numeric(MTurk_education_15[5,2]) + as.numeric(MTurk_education_15[7,2])
MTurk_education_15 <- rbind(no_highschool_dip, MTurk_education_15, advanced_deg_15)
MTurk_education_15[10,1] <- c("bachelor's or higher")
MTurk_education_15[1,1] <- c("less than a high school diploma")
MTurk_edu_15 <- MTurk_education_15[c(-2, -4, -6, -7, -8, -9), ]
MTurk_edu_15 <- MTurk_edu_15[c(1, 3, 2, 4), ]

#Define confidence intervals 
eduCI_15 <- binom.confint(x = MTurk_edu_15$n, n = sum(MTurk_edu_15$n), conf.level = 0/95, methods = "exact")
eduCI_15 <- eduCI_15[ , 4:6]

#Build Education table 
education_table_15 <- cbind(MTurk_edu_15, eduCI_15, BLS_edu_15)
names(education_table_15) <- c("education level", "n", "o", "LL_eth", "UL_eth", "e")
education_table_15 <- education_table_15[ , c(1, 6, 2, 4, 3, 5)]

#Education chi squared test
edu_ch2_15 <- as.data.frame(chi2(education_table_15))
names(edu_ch2_15) <- c("chi2")
row.names(edu_ch2_15) <- c("less than a highschool diploma","high school/GED" , "2 year degree", "bachelor's or higher")
edu_ch2_15 <- na.omit(edu_ch2_15)

#Sum chi-square values.
sum(edu_ch2_15$chi2, na.rm=TRUE)
```

##Study 2 Age
```{r}
#Rachel's 2015 Demographic Analysis 
library(readxl)
MTurk15_R <- read_excel("Rachel's Data from 2015.xlsx")

#packages to clean data for demogrphics
library(stringr)
install.packages(tidyr)
library(tidyr)

#Rachel's age analysis 
#Load BLS Data
BLS_age_15 <- read_excel("~/Desktop/BLS.Age.2015.xlsx")

#Define population mean. In this case it was already calculated in the dataset. 
pop_mean_15 <- as.numeric(BLS_age_15[6, 10])

#Define sample mean
samp_mean_15_R <- as.numeric(MTurk15_R %>% 
                             summarise(MTurk_age = mean(Age, na.rm = TRUE)))

#Sample standard deviation.
samp_sd_15_R <- as.numeric(MTurk15_R %>% 
                           summarise(sd_age = sd(Age, na.rm = TRUE)))

#Define sample size
n_15_R <- nrow(MTurk15_R)

#Run t test.
t_value_15_R <- (samp_mean_15_R - pop_mean_15)/(samp_sd_15_R/sqrt(n_15_R))
```

##Study 2 Gender
```{r}
#Rachel's gender analysis
#clean data
MTurk15_R$Gender <- str_replace(MTurk15_R$Gender, "1", "male")
MTurk15_R$Gender <- str_replace(MTurk15_R$Gender, "2", "female")

#get frequencies
MTurk_gender_15_R <- MTurk15_R %>% 
  count(Gender)
MTurk_gender_15_R <- MTurk_gender_15_R[1:2,]

#Define confidence intervals
require(binom)
gCI_15_R <- binom.confint(x = MTurk_gender_15_R$n, n = sum(MTurk_gender_15_R$n), conf.level = 0.95, methods = "exact")
gCI_15_R <- gCI_15_R[ ,4:6]

#Build Gender table with gender, expected value, n, LL, observed, UL
gender_table_15_R <- cbind(MTurk_gender_15_R, gCI_15_R, BLS_gender_15)
names(gender_table_15_R) <- c("gender", "n", "o", "LL_g", "UL_g", "e")
gender_table_15_R <- gender_table_15_R[, c(1, 3, 2, 4, 6, 5)]

#Gender chi square test
gender_ch2_15_R <- as.data.frame(chi2(gender_table_15_R))
names(gender_ch2_15_R) <- c("chi2")
gender_ch2_15_R <- na.omit(gender_ch2_15_R)

#Sum chi-square values.
sum(gender_ch2_15_R$chi2, na.rm=TRUE)

##Study 2 Ethnicity
```{r}
#Rachel's ethnicity analysis
MTurk15_R_white <- length(which(MTurk15_R$White == 1))
MTurk15_R_AfnAmn <- length(which(MTurk15_R$AfnAmn == 1))
MTurk15_R_AsnPI <- length(which(MTurk15_R$AsnPI == 1))

MTurk15_R_eth <- as.data.frame(c(MTurk15_R_white, MTurk15_R_AfnAmn, MTurk15_R_AsnPI))
MTurk15_R_eth <- cbind(BLS_ethnicity_15_data$ethnicity, MTurk15_R_eth)
colnames(MTurk15_R_eth) <- c("ethnicity", "n")
MTurk15_R_eth_tot <- MTurk15_R_white + MTurk15_R_AfnAmn + MTurk15_R_AsnPI

#Get proportions for ethnicities
MTurk15_R_eth <- MTurk15_R_eth %>% 
  mutate(prop = sapply(MTurk15_R_eth$n, function(x) {x/MTurk15_R_eth_tot}))

#Define confidence intervals 
ethCI_15_R <- binom.confint(x = MTurk15_R_eth$n, n = sum(MTurk15_R_eth$n), conf.level = 0/95, methods = "exact")
ethCI_15_R <- ethCI_15_R[ , 4:6]

#Build Ethnicity table 
ethnicity_table_15_R <- cbind(MTurk15_R_eth[1:2], ethCI_15_R, BLS15_eth_prop)
names(ethnicity_table_15_R) <- c("ethnicity", "n", "o", "LL", "UL", "e")
eth_table_15_R <- ethnicity_table_15_R[1:3, c(1, 6, 2, 4, 3, 5)]

#Ethnicity chi squared test
eth_ch2_15_R <- as.data.frame(chi2(eth_table_15_R))
names(eth_ch2_15_R) <- c("chi2")
row.names(eth_ch2_15_R) <- c(BLS_ethnicity_15_data$ethnicity)
eth_ch2_15_R <- na.omit(eth_ch2_15_R)

#Sum chi-square values.
sum(eth_ch2_15_R$chi2, na.rm=TRUE)
```

##Study 2 Education Level 
```{r}
#Rachel's education level
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "1", "Less than high school")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "2", "High school/GED")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "3", "Some college")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "4", "2 year deg")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "5", "4 year deg")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "6", "Master's")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "7", "Doctoral")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "8", "J.D.")
MTurk15_R$Educ <- str_replace(MTurk15_R$Educ, "9", "M.D.")

MTurk15_R_edu <- MTurk15_R %>%
  count(Educ)

no_highschool_dip_R <- length(which(MTurk15_R$Educ == "Less than high school"))
highschool_ged_R <- length(which(MTurk15_R$Educ == "High school/GED", MTurk15_R$Educ == "Some college"))
two_year_deg_R <- length(which(MTurk15_R$Educ == "2 year deg"))
advanged_deg_15_R <- length(which(MTurk15_R$Educ == "4 year deg")) + length(which(MTurk15_R$Educ == "Master's")) + length(which(MTurk15_R$Educ == "Doctoral")) + length(which(MTurk15_R$Educ == "J.D.")) + length(which(MTurk15_R$Educ == "M.D."))

MTurk15_R_edu <- as.data.frame(c(no_highschool_dip_R, highschool_ged, two_year_deg_R, advanged_deg_15_R))
education_levels <- c("no high school", "high school dimploma/GED", "2 year degree", "Bachelors or higher")
MTurk15_R_edu <- cbind(education_levels, MTurk15_R_edu)
colnames(MTurk15_R_edu) <- c("education level", "n")

#Define confidence intervals 
eduCI_15_R <- binom.confint(x = MTurk15_R_edu$n, n = sum(MTurk15_R_edu$n), conf.level = 0/95, methods = "exact")
eduCI_15_R <- eduCI_15_R[ , 4:6]

#Build Education table 
education_table_15_R <- cbind(MTurk15_R_edu, eduCI_15_R, BLS_edu_15)
names(education_table_15_R) <- c("education level", "n", "o", "LL_eth", "UL_eth", "e")
education_table_15_R <- education_table_15_R[ , c(1, 6, 2, 4, 3, 5)]

#Education chi squared test
edu_ch2_15_R <- as.data.frame(chi2(education_table_15_R))
names(edu_ch2_15_R) <- c("chi2")
row.names(edu_ch2_15_R) <- c("less than a highschool diploma","high school/GED" , "2 year degree", "bachelor's or higher")
edu_ch2_15_R <- na.omit(edu_ch2_15_R)

#Sum chi-square values.
sum(edu_ch2_15_R$chi2, na.rm=TRUE)
```

##Replication Demographic Analaysis
##2017 Occupational Analysis 
```{r}
#Load datasets. Note: these datasets were not taken direclty from their respective sources (i.e., BLS or Qualtrics survey, respectively). The BLS dataset is a table describing the broader occupational categories, number of employees working in jobs that fall under one of these categories, and the relative frequency of jobs falling under said occupational category. The LATech dataset comes from a separate investigaiton using MTurk whereby individuals who were flagged for inattentive responding were removed from the dataset. 
#Build the BLS Table, but first establish a temp file.

##BLS Data
library(readxl)
BLS <- read_excel("2017 BLS Data-updated.xlsx")
BLS <- BLS[c(8,10)]

#Include data on the size of the military (taken from (https://www.bls.gov/ooh/military/military-careers.htm))
BLS <- rbind(BLS, c("Military specific occupations", 1085286))

#Only major occupation groups are included in BLS
BLS <- 
  BLS[c(2,71,127,159,219,290,316,332,428,490,581,607,648,682,700,760,803,916,943,1046,1122,1288,427267,1),]

#Add military occupation number to the overall
BLS$Var2[14] <- BLS$Var2[13] + BLS$Var2[14]

#Define new BLS. "e" is for "relative frequency". 
BLS <- tab[c(1,3)]
names(BLS) <- c("Intermediate Aggregation Title", "e")

##LATech data
LATech <- read_excel("2017 Data.xls")

###Delete unneeded data.
LATech <- LATech[c(12, 20:21)]
names(LATech) <- c("Self-Reported Job Title", "O*NET Occupation Title", "O*NET Job Code")

###Edit data. Some participants provided incorrect codes or no codes.
LATech[644, 3] = "41-9031.00"
LATech[274,] = c("analyst", "analyst", "13-0000.00")

###Remove unemployed cases.
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}
LATech <- completeFun(LATech, "O*NET Job Code")

###Build relative frequency table.
tab <- as.data.frame(table(LO$Int.Agg.Cat, exclude=NULL))
tab <- transform(tab, relative = prop.table(Freq))
names(tab) <- c("Intermediate Aggregation Title","n","f")
tab <- merge(tab,BLS, by="Intermediate Aggregation Title", all=TRUE)

#Confidence interval formula for proportions with assymptotic intervals. Intervals are assymptotic because proportions cannot be less than 0 or greater than 1, which becomes problematic as proportions near 0 or 1. 
require("binom")
CI <- binom.confint(tab$n, sum(tab$n), conf.level = 0.95, methods = "exact")
CI <- CI[c(5:6)]
table <- cbind(tab,CI)
table <- table[,c(1,4,2,5,3,6)]
ch <- as.data.frame(chi2(table))
ch <- na.omit(ch)
names(table) <- c("Intermediate Aggregation Title","Exp(p)","n1","LCL1","Obs(p)1","UCL1")

#Sum chi-square values.
sum(ch$`chi2(table)`, na.rm=TRUE)

##Validate using chi-square function in r.
#Function for differences in relative percentages. This is not printed in the tables, but could be used to help readers see the disparity between expected and observed proportions.
f <- function(x) {
  f   <- (x$f)
  BLS <- (x$e)
  (f-BLS)
}
f1 <- as.data.frame(f(table))

#Print chi-square APA style.
require(devtools)
install_github("dgromer/apa")
require(apa)
##Sample 1#
observed1 <- table$n1
observed1 <- observed1[c(1:13)]
expected1 <- table$`Exp(p)`
expected1 <- expected1[c(1:13)]
c1 <- chisq_apa(chisq.test(x = observed1, p = expected1, correct = TRUE))

#Post-hoc power analysis.
##Convert chi-squares to correlation coefficient to allow meaningful statistic.
library(pwr)
###Study 1
X21 <-sum(ch$`chi2(table)`, na.rm=TRUE)
r1 <- sqrt((X21/sum(tab$n)))
pwr.chisq.test(w=r1,N=(sum(tab$n)),df=12,sig.level = .01)
```

##2017 Age Analysis
```{r}
#Age Analysis T- Test
#Load Dataset
require(dplyr)
require(readxl)
age_demo <- read_excel("age_demo.xlsx")

@@

##2017 Gender Analysis
```{r}
##Gender Analysis
BLS_data <- read_excel("employment_demo.xlsx")

@@

##2017 Ethnicity Analysis
```{r}
#Ethnicity Analysis 
#Define obs_eth
BLS_cauc <- as.numeric(BLS_data[7, 4])/100
BLS_AfAm <- as.numeric(BLS_data[7,5])/100
BLS_Asn <- as.numeric(BLS_data[7,6])/100
BLS_Hisp <- as.numeric(BLS_data[7,7])/100
BLS_eth <- c(BLS_AfAm, BLS_Asn, BLS_cauc, BLS_Hisp)

@@

##2017 Education Analysis 
```{r}
#Education Analysis 
#Clean BLS Data. Get totals.  
BLS_education_original <- read_excel("education_demo.xlsx")
BLS_education <- as.data.frame(BLS_education_original[8:14, ])
BLS_education <- BLS_education[1, c(-2, -3, -6)]
names(BLS_education) <- c("category", "High school graduates, no college(1)", "Some college, no degree", "Associate degree", "Bachelor's degree only", "Advanced degree")

@@

#Use vapply and the above function to get expected proportions from BLS data
BLS_edu <- sapply(BLS_education, function(x) {x/BLS_total})
